---
title: 'Linear methods: principal components regression'
author: "Jaewoo Cho"
date: '2023-02-13'
output: github_document
---
# Use the prostate cancer data.
```{r}
library('splines')        ## for 'bs'
library('dplyr')          ## for 'select', 'filter', and others
library('magrittr')       ## for '%<>%' operator
library('glmnet')         ## for 'glmnet'

prostate <- read.table(url('https://web.stanford.edu/~hastie/ElemStatLearn/datasets/prostate.data'))

pairs(prostate)

## split prostate into testing and training subsets
prostate_train <- prostate %>%
  filter(train == TRUE) %>% 
  select(-train)

prostate_test <- prostate %>%
  filter(train == FALSE) %>% 
  select(-train)

```
# Use the cor function to reproduce the correlations listed in HTF Table 3.1, page 50.
```{r}
cor(prostate)
```


# Treat lcavol as the outcome, and use all other variables in the data set as predictors.
```{r}
## predict lcavol consider all other predictors
## lm fits using L2 loss
fit <- lm(lcavol ~ ., data=prostate_train)
summary(fit)

```

# With the training subset of the prostate data, train a least-squares regression model with all predictors using the lm function.
```{r}
fit <- lm(lcavol ~ ., data=prostate_train)
coef(fit)
residuals(fit)
```

#Use the testing subset to compute the test error (average squared-error loss) using the fitted least-squares regression model.
```{r}
## functions to compute testing/training error w/lm
L2_loss <- function(y, yhat)
  (y-yhat)^2
error <- function(dat, fit, loss=L2_loss)
  mean(loss(dat$lcavol, predict(fit, newdata=dat)))

## train_error 
error(prostate_train, fit)

## testing error
error(prostate_test, fit)
```

#Train a ridge regression model using the glmnet function, and tune the value of lambda (i.e., use guess and check to find the value of lambda that approximately minimizes the test error).
```{r}

## use glmnet to fit lasso
## glmnet fits using penalized L2 loss
## first create an input matrix and output vector
form  <- lcavol ~  lweight + age + lbph + lcp + pgg45 + lpsa + svi + gleason
x_inp <- model.matrix(form, data=prostate_train)
y_out <- prostate_train$lcavol
fit <- glmnet(x=x_inp, y=y_out, lambda=seq(0.5, 0, -0.05))
print(fit$beta)

## functions to compute testing/training error with glmnet
error <- function(dat, fit, lam, form, loss=L2_loss) {
  x_inp <- model.matrix(form, data=dat)
  y_out <- dat$lcavol
  y_hat <- predict(fit, newx=x_inp, s=lam)  ## see predict.elnet
  mean(loss(y_out, y_hat))
}

## train_error at lambda=0
error(prostate_train, fit, lam=0, form=form)

## testing error at lambda=0
error(prostate_test, fit, lam=0, form=form)

## train_error at lambda=0.03
error(prostate_train, fit, lam=0.05, form=form)

## testing error at lambda=0.03
error(prostate_test, fit, lam=0.05, form=form)
```

# Create a figure that shows the training and test error associated with ridge regression as a function of lambda
```{r}

## compute training and testing errors as function of lambda
err_train_1 <- sapply(fit$lambda, function(lam) 
  error(prostate_train, fit, lam, form))
err_test_1 <- sapply(fit$lambda, function(lam) 
  error(prostate_test, fit, lam, form))

## plot test/train error
plot(x=range(fit$lambda),
     y=range(c(err_train_1, err_test_1)),
     xlim=rev(range(fit$lambda)),
     type='n',
     xlab=expression(lambda),
     ylab='train/test error')
points(fit$lambda, err_train_1, pch=19, type='b', col='darkblue')
points(fit$lambda, err_test_1, pch=19, type='b', col='darkred')
legend('topright', c('train','test'), lty=1, pch=19,
       col=c('darkblue','darkred'), bty='n')

colnames(fit$beta) <- paste('lam =', fit$lambda)
#print(fit$beta %>% as.matrix)
```

# Create a path diagram of the ridge regression analysis, similar to HTF Figure 3.8
```{r}
## use glmnet to fit lasso
## glmnet fits using penalized L2 loss
## first create an input matrix and output vector
form  <- lcavol ~  lweight + age + lbph + lcp + pgg45 + lpsa + svi + gleason
x_inp <- model.matrix(form, data=prostate_train)
y_out <- prostate_train$lcavol
fit <- glmnet(x=x_inp, y=y_out, lambda=seq(0.5, 0, -0.05))
print(fit$beta)

## plot path diagram
plot(x=range(fit$lambda),
     y=range(as.matrix(fit$beta)),
     type='n',
     xlab=expression(lambda),
     ylab='Coefficients')
for(i in 1:nrow(fit$beta)) {
  points(x=fit$lambda, y=fit$beta[i,], pch=19, col='blue')
  lines(x=fit$lambda, y=fit$beta[i,], col='blue')
}
text(x=0, y=fit$beta[,ncol(fit$beta)], 
     labels=rownames(fit$beta),
     xpd=NA, pos=4, srt=45)
abline(h=0, lty=3, lwd=2)


```


