---
title: "Homework 1	Least-squares, nearest-neighbors"
date: January 15, 2020
author: Jaewoo Cho
output: github_document
---
## Libraries
```{r}
library('class')
library('dplyr')
```
# 1. Paste the code from the mixture-data-lin-knn.R file into the homework template Knitr document.
## Source Code

```{r}
## load binary classification example data from author website 
## 'ElemStatLearn' package no longer available
load(url('https://web.stanford.edu/~hastie/ElemStatLearn/datasets/ESL.mixture.rda'))
dat <- ESL.mixture

plot_mix_data <- function(dat, datboot=NULL) {
  if(!is.null(datboot)) {
    dat$x <- datboot$x
    dat$y <- datboot$y
  }
  plot(dat$x[,1], dat$x[,2],
       col=ifelse(dat$y==0, 'blue', 'orange'),
       pch=20,
       xlab=expression(x[1]),
       ylab=expression(x[2]))
  ## draw Bayes (True) classification boundary
  prob <- matrix(dat$prob, length(dat$px1), length(dat$px2))
  cont <- contourLines(dat$px1, dat$px2, prob, levels=0.5)
  rslt <- sapply(cont, lines, col='purple')
}
plot_mix_data(dat)
```
```{r}
## fit linear classifier
fit_lc <- function(y, x) {
  x <- cbind(1, x)
  beta <- drop(solve(t(x)%*%x)%*%t(x)%*%y)
}

## make predictions from linear classifier
predict_lc <- function(x, beta) {
  cbind(1, x) %*% beta
}

## fit model to mixture data and make predictions
lc_beta <- fit_lc(dat$y, dat$x)
lc_pred <- predict_lc(dat$xnew, lc_beta)

## reshape predictions as a matrix
lc_pred <- matrix(lc_pred, length(dat$px1), length(dat$px2))
contour(lc_pred,
      xlab=expression(x[1]),
      ylab=expression(x[2]))

```
```{r}
## find the contours in 2D space such that lc_pred == 0.5
lc_cont <- contourLines(dat$px1, dat$px2, lc_pred, levels=0.5)

## plot data and decision surface
plot_mix_data(dat)
sapply(lc_cont, lines)
```

```{r}
## fit knn classifier
## use 5-NN to estimate probability of class assignment
knn_fit <- knn(train=dat$x, test=dat$xnew, cl=dat$y, k=5, prob=TRUE)
knn_pred <- attr(knn_fit, 'prob')
knn_pred <- ifelse(knn_fit == 1, knn_pred, 1-knn_pred)

## reshape predictions as a matrix
knn_pred <- matrix(knn_pred, length(dat$px1), length(dat$px2))
contour(knn_pred,
        xlab=expression(x[1]),
        ylab=expression(x[2]),
        levels=c(0.25, 0.5, 0.75))
```


```{r}
## find the contours in 2D space such that knn_pred == 0.5
knn_cont <- contourLines(dat$px1, dat$px2, knn_pred, levels=0.5)

## plot data and decision surface
plot_mix_data(dat)
sapply(knn_cont, lines)
```
```{r}
 ## do bootstrap to get a sense of variance in decision surface
resample <- function(dat) {
  idx <- sample(1:length(dat$y), replace = T)
  dat$y <- dat$y[idx]
  dat$x <- dat$x[idx,]
  return(dat)
}

## plot linear classifier for three bootstraps
par(mfrow=c(1,3))
for(b in 1:3) {
  datb <- resample(dat)
  ## fit model to mixture data and make predictions
  lc_beta <- fit_lc(datb$y, datb$x)
  lc_pred <- predict_lc(datb$xnew, lc_beta)
  
  ## reshape predictions as a matrix
  lc_pred <- matrix(lc_pred, length(datb$px1), length(datb$px2))

  ## find the contours in 2D space such that lc_pred == 0.5
  lc_cont <- contourLines(datb$px1, datb$px2, lc_pred, levels=0.5)
  
  ## plot data and decision surface
  plot_mix_data(dat, datb)
  sapply(lc_cont, lines)
}
```


```{r}
## plot 5-NN classifier for three bootstraps
par(mfrow=c(1,3))
for(b in 1:3) {
  datb <- resample(dat)
  
  knn_fit <- knn(train=datb$x, test=datb$xnew, cl=datb$y, k=5, prob=TRUE)
  knn_pred <- attr(knn_fit, 'prob')
  knn_pred <- ifelse(knn_fit == 1, knn_pred, 1-knn_pred)
  
  ## reshape predictions as a matrix
  knn_pred <- matrix(knn_pred, length(datb$px1), length(datb$px2))

  ## find the contours in 2D space such that knn_pred == 0.5
  knn_cont <- contourLines(datb$px1, datb$px2, knn_pred, levels=0.5)
  
  ## plot data and decision surface
  plot_mix_data(dat, datb)
  sapply(knn_cont, lines)
}

```


```{r}
## plot 20-NN classifier for three bootstraps
par(mfrow=c(1,3))
for(b in 1:3) {
  datb <- resample(dat)
  
  knn_fit <- knn(train=datb$x, test=datb$xnew, cl=datb$y, k=20, prob=TRUE)
  knn_pred <- attr(knn_fit, 'prob')
  knn_pred <- ifelse(knn_fit == 1, knn_pred, 1-knn_pred)
  
  ## reshape predictions as a matrix
  knn_pred <- matrix(knn_pred, length(datb$px1), length(datb$px2))
  
  ## find the contours in 2D space such that knn_pred == 0.5
  knn_cont <- contourLines(datb$px1, datb$px2, knn_pred, levels=0.5)
  
  ## plot data and decision surface
  plot_mix_data(dat, datb)
  sapply(knn_cont, lines)
}
```

# 2. Read the help file for R's built-in linear regression function lm
- lm is used to fit linear models. It can be used to carry out regression, single stratum analysis of variance and analysis of covariance (although aov may provide a more convenient interface for these).
- lm(formula, data, subset, weights, na.action, method = "qr", model = TRUE, x = FALSE, y = FALSE, qr = TRUE, singular.ok = TRUE, contrasts = NULL, offset, ...)
```{r}
?lm
```

# 3. Re-write the functions fit_lc and predict_lc using lm, and the associated predict method for lm objects.

```{r}
## fit linear classifier
fit_lc <- function(y, x) {
  dat <- data.frame(y=y, x1=x[,1], x2 = x[,2])
  fit <- lm(y ~ x1 + x2, data = dat)
  fit
}

## make predictions from linear classifier
predict_lc <- function(x, fit) {
  predict(fit, data.frame(x))
}

## fit model to mixture data and make predictions
lc_beta <- fit_lc(dat$y, dat$x)
lc_pred <- predict_lc(dat$xnew, lc_beta)

## reshape predictions as a matrix
lc_pred <- matrix(lc_pred, length(dat$px1), length(dat$px2))
contour(lc_pred,
      xlab=expression(x[1]),
      ylab=expression(x[2]))
## find the contours in 2D space such that knn_pred == 0.5
knn_cont <- contourLines(dat$px1, dat$px2, knn_pred, levels=0.5)

## plot data and decision surface
plot_mix_data(dat)
sapply(knn_cont, lines)
```

# 4. Consider making the linear classifier more flexible, by adding squared terms for x1 and x2 to the linear model

```{r}
## fit linear classifier
fit_lc <- function(y, x) {
  dat <- data.frame(y=y, x1=x[,1], x2 = x[,2])
  fit <- lm(y ~ (x1^2) + (x2^2), data = dat)
  fit
}

## make predictions from linear classifier
predict_lc <- function(x, fit) {
  predict(fit, data.frame(x))
}

## fit model to mixture data and make predictions
lc_beta <- fit_lc(dat$y, dat$x)
lc_pred <- predict_lc(dat$xnew, lc_beta)

## reshape predictions as a matrix
lc_pred <- matrix(lc_pred, length(dat$px1), length(dat$px2))
contour(lc_pred,
      xlab=expression(x[1]),
      ylab=expression(x[2]))
## find the contours in 2D space such that knn_pred == 0.5
knn_cont <- contourLines(dat$px1, dat$px2, knn_pred, levels=0.5)

## plot data and decision surface
plot_mix_data(dat)
sapply(knn_cont, lines)
```

# 5. Describe how this more flexible model affects the bias-variance tradeoff

- The flexible model affects the bias-variance tradeoff by having less variance within the model, which leads to the increase in bias for the model. This happens vice versal for models that are less flexible models with the tradeoff of more variance and less bias.




